### 2008-09-19 doc/papers/cgo08/paper.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@349 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

% CGO 09: http://www.cgo.org/
% svn co https://svn.mcs.anl.gov/repos/performance/orio/doc/papers/cgo08
% Max 22 pages, double-spaced, 11pt or larger font
% Abstract 300 words: Sep. 12, 2008
% Full paper: Sep. 19, 2008

\documentclass[11pt]{article}

\usepackage{url}
\usepackage{html}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}

\title{Generating Empirically Optimized Numerical Software from MATLAB Prototypes}

\author{
Elizabeth Jessup,$^1$ Boyana Norris,$^2$ Albert Hartono,$^3$ and Jeremy Siek$^4$\\
\\
\vspace{-.05in}
{\small $^1$University of Colorado at Boulder Dept. of Computer Science, Boulder, CO~80309}\\
\vspace{-.05in}
{\small \url{jessup@cs.colorado.edu}}\\
\vspace{-.05in}
{\small $^2$Argonne National Laboratory Mathematics and Computer Science Division, Argonne, IL~60439}\\
\vspace{-.05in}
{\small \url{norris@mcs.anl.gov}}\\
\vspace{-.05in}
{\small $^3$Ohio State University Dept. of Computer Science and Engineering, Columbus, OH~43210}\\
\vspace{-.05in}
{\small \url{hartonoa@cse.ohio-state.edu}}\\
\vspace{-.05in}
{\small $^4$University of Colorado at Boulder Dept. of Electrical and Computer Engineering, Boulder, CO~80309}\\
\vspace{-.05in}
{\small \url{jeremy.siek@colorado.edu}}
}

\begin{document}

\maketitle

\abstract{

The growing demand for higher levels of detail and accuracy in results means that the size and complexity of scientific computations is increasing at least as fast as the improvements in processor technology. Programming scientific applications is hard, and optimizing them for high performance is even harder.  The development of optimized codes requires extensive knowledge, not only of the costs of floating-point arithmetic but also of  memory access issues and compiler optimizations. Experiments show that the complexity of this hardware-software system means that performance is difficult to predict fully.  Therefore, computational scientists are often forced to choose between investing too much time in tuning code or  accepting performance that is significantly lower than the best achievable performance on a given architecture.

In this paper, we describe the first steps toward a fully automated system for the optimization of the matrix algebra kernels that are a foundational part of many scientific applications.  To generate highly optimized code from a high-level MATLAB prototype, we will follow a three-step approach.  To begin, we have developed a compiler that converts a MATLAB script into simple C code.   We then use the source transformation tool PLuTo to optimize that code for coarse-grained parallelism and locality simultaneously. The result is an automatically generated code containing loop transformations and OpenMP annotations for exploiting multicore parallelism. Finally, we employ the empirical performance tuning system Orio to generate many tuned versions of the same operation using different optimization techniques, such as loop unrolling and memory alignment. Orio performs an automated empirical search for selecting the best among the multiple optimized code variants.

We discuss performance results showing that the code generated by using our system significantly outperforms not only the original simple C code but also code based on source BLAS, ATLAS-optimized BLAS, and Intel MKL routines.
}

\doublespacing
\section{Introduction}
\label{sec:intro}

The development of high-performance numerical codes is challenging because performance is determined by complex interactions between the algorithm, data structure, programming language, compiler, and computer architecture. Scientists seeking high performance are thus required to master advanced concepts in computer science and carry out intricate programming tasks in addition to managing the scientific content of their work.  They must either invest substantial time in tuning their software or accept low performance. In either case, the productivity of the scientist degrades.

Historically, the research community has pursued two separate paths toward the goal of making software run at near-peak levels.  The first of these builds on research into compilers and their associated technologies.  The holy grail of compilation research is to take an arbitrary code as input and produce optimal code as output for a given language and hardware platform.  Despite the immense effort that has been poured into that approach, its success has been limited both by practical constraints (users will not tolerate compile times that extend into several days) and by the amount of detailed information the compiler can obtain about the software to be compiled and the hardware on which it is to execute~\cite{Ken99,Wol89,Wol96}.    The second path has been to identify kernel routines that constitute the dominant costs of a wide variety of applications.  When such kernels can be identified and agreed upon by the members of the community, programmers with the required level of technical knowledge can concentrate on producing these optimized kernel libraries for architectures of interest. A prime example is the high-performance BLAS libraries produced by a combination of hardware vendors, independent software vendors, and researchers.  Developers who write their codes calling these routines can achieve high performance across all supported architectures. The efforts in high-performance library development aptly demonstrate, however, that the research into compiler technologies has not been fully successful.  If compilers could take the relatively simple code even for basic tasks such as a dense matrix-matrix product and produce optimal code, this second approach would be unnecessary.    Like compiler optimization, the kernel library-oriented approach has significant limitations.  Support for a variety of architectures is the most serious problem because the optimizations necessary to achieve near-peak performance are by nature not portable.   Lack of portability is particularly problematic because processor designs are constantly changing,  and constant updates are nearly impossible to maintain by hand.  It is a sad fact of the computing industry that by the time highly-optimized code is available for a given architecture, that architecture is generally well on its way to obsolescence.

We are working on a combination of the two approaches designed to overcome some of their shortcomings.   This paper described our three-pronged approach toward the development of an automatically tuned library for matrix algebra computations.  In Section~\ref{sec:matlab}, we describe the compiler we have developed that converts a MATLAB script into simple C code. In Section~\ref{sec:orio}, we demonstrate how to employ the empirical performance tuning system Orio in conjunction with the PLuTo tool to generate many tuned versions of the same operation using different optimization techniques, such as loop unrolling and memory alignment.  The result is an automatically generated code containing loop transformations and OpenMP annotations for exploiting multicore parallelism. In Section~\ref{sec:experiments}, we provide performance results showing that the code generated by using our system significantly outperforms not only the original simple C code but also code based on source BLAS, ATLAS-optimized BLAS, and Intel MKL routines.
%Text in intro repeats abstract some, but that might not be a big deal for a conf paper.

\section{Background}
\label{sec:background}

%Modified from recent DOE proposal

The most common approach  to tuning numerical codes is for an expert to transform the source manually, unrolling loops, blocking for multiple levels of cache, and inserting prefetch instructions.  The pitfalls of this approach are well-understood~\cite{Goedecker01}.  It requires a significant amount of time and effort. Optimizing code for one particular platform may in fact make it less efficient on other platforms and often makes it complex and hard to understand or maintain.    An alternative is the use of tuned libraries of key numerical algorithms (e.g., BLAS~\cite{Dongarra:1990fk} and LAPACK~\cite{LAPACK} for dense linear algebra, SPARSPAK~\cite{sparspak} for sparse linear algebra, and BLITZ++~\cite{blitz} to support scientific computing in C++, or GPULib by Tech-X, an open source library for accelerating numerical computations on Graphics Processing Units~\cite{gpulib}, to name just a few). Such libraries are difficult to develop, and they generally cannot adapt to execution context: as we note in Section~\ref{sec:composed}, a series of calls to library routines can incur high memory costs.

Specialized code generators circumvent the high costs of manual code generation . They include tools for basic dense linear algebra operations (ATLAS~\cite{WN147}, PhiPAC~\cite{bilmes97optimizing,phipacwww}), sparse linear algebra (OSKI~\cite{vuduc05}), FFTs~\cite{FFTW,Spiral}), stencil-based operations~\cite{kamil06}, and tensor computations (TCE~\cite{TCE}) among others.  It is often impossible to predict the performance of code precisely on modern computer architectures. Thus, many of these specialized code generators exploit search strategies to identify the best (or nearly best) code for a particular choice of problem parameters and machine. Such autotuning tools are not general but rather each is valuable in one specific problem domain. While not all applications can be written in terms of available libraries, the performance achieved with autotuning tools does convincingly demonstrate the advantage of source-to-source transformation over optimizing compilers.

The best performance today is obtained through approaches like that taken by PERI~\cite{PERI} which involves a labor intensive rewrite of portion of the code by a group of performance tools experts in collaboration with application scientists. Maintaining performance of the code as the application and hardware evolves means that the tuning process may need to be repeated. We propose to investigate ways to capture this performance tuning expertise in a way that can be integrated into the application. We will thus enable performance portability and ensure that the improvements to the application are long-lasting.

A number of source or binary transformation tools for performance-improving optimizations exist. LoopTool~\cite{LoopTool}, developed at Rice University, supports annotation-based loop fusion, unroll-and-jam, skewing and tiling.  A new tool, POET~\cite{POET} also supports a number of loop transformations. POET offers a complex template-based syntax for defining transformations in a language-independent manner. Other recent research efforts whose goal, at least in part, is to enable optimizations of source code to be augmented with performance-related information include the X language~\cite{XLanguage} (a macro C-like language for annotating C code), the Broadway~\cite{broadway} compiler,  telescoping languages~\cite{telescopingurl,teleoverview,Ken99}, and various meta-programming techniques~\cite{veldhuizen95,weise93,kiczales91,chiba95}. PLuTo~\cite{Pluto,uday08cc} is a source-to-source transformation tool for optimizing sequences of nested loops. PLuTo employs a polyhedral model of nested loops, where the dynamic instance (iteration) of each statement is viewed as an integer point in a well-defined space, called the statement's polyhedron. Combined with a characterization of data dependences, this representation allows the construction of mathematically correct complex loop transformations. The transformations target both improved cache locality and parallelism.

\section{Composed Operations}
\label{sec:composed}

Codes based on matrix algebra are generally constructed as a sequence of calls to the Basic Linear Algebra Subprograms (BLAS)~\cite{Dongarra:1988uq,Lawson:1979kx,Dongarra:1990fk} and similar sparse matrix libraries~\cite{Saad:fr,George:1981uq}. Writing programs in this way promotes readability and maintainability but can be costly in terms of memory efficiency. Specifically, the retrieval of a large order matrix at each routine call can have a profound effect on performance even when highly tuned implementations of the BLAS (e.g ~\cite{Bilmes:1997ye,Whaley:1998fk,IntelMath:oq,ESSL:kl,Goto:2006fk}) are used.

The authors and other researchers have observed that it can be much more efficient to call a single specialized routine that performs multiple operations than to make successive calls to separate BLAS ~\cite{Ashby:uq,Blackford:2002vn,Stanley:CSD-98-992,baker03blgmres,baker03lgmres,Dennis:2005tg,Howell:2008,gropp01,Vuduc:2003kl}.  Single routines that carry out more than one linear algebra operation are known as {\t composed} BLAS. As an example, consider the pair of matrix-vector products $q  = Ap$, $s  = A^Tr$, where $A$ is a matrix and $p$, $q$, $r$, and $s$ are vectors, that comprise the computational
bottlenecks of the BiConjugate Gradient Method~\cite{Barrett:1994kx,Saad:2003fk}  and of the GEMVER kernel examined in Section~\ref{sec:matlab}.   These two operations can be implemented as a pair of calls to the BLAS routine {\tt GEMV} or they can be rewritten as a composed BLAS consisting of the doubly nested loop encompassing both matrix-vector multiplies. In the former case, the matrix $A$ is accessed twice while, in the latter, is it accessed only once. Our preliminary results indicate that this loop fusion leads to a routine that delivers 90\% more megaflops than does a pair of calls to the highly optimized Goto BLAS~\cite{Goto:2006fk} {\tt GEMV} routines for large matrix orders.

Composed routines are the focus of the work presented here, but much of what we discuss generalizes to a much broader array of computations.   Extending our approach will be the subject of future work.

% While large speedups are probable, composition can  also force memory accesses to be consecutive, and it can

%lead to the need to bring more data into cache. These bad

% effects can be remedied by combining composition with other tuning techniques like cache blocking and software pipelining.

\section{Our Approach}
\label{sec:approach}

To generate highly optimized code from a MATLAB prototype, we will follow a three-step approach, illustrated in Figure~\ref{fig:process}.    To begin, we have developed a compiler that converts a MATLAB script into simple C code.   We then use the source-to-source automatic parallelization tool  PLuTo\cite{Pluto} to optimize for coarse-grained parallelism and locality simultaneously. The result is an automatically generated code containing transformed loops and OpenMP directives for compiler parallelization.  Finally, we employ the extensible annotation system Orio to generate many tuned versions of the same operation using different optimization parameters and perform an empirical search for selecting the best among the multiple optimized code variants.
%Does this repeat abstract/intro too closely?

%Probably need some words about WHY this 3 pronged approach is a good idea.

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/process.png}

\caption{Overview of the code generation and tuning process.}
\label{fig:process}
\end{figure}

\subsection{A MATLAB Compiler}
\label{sec:matlab}

We have developed a compiler that translates MATLAB scripts into C code.    Figure~\ref{fig:compiler} gives an overview of the compilation process. The MATLAB kernel specification is parsed into a high-level intermediate representation in the form of a dataflow graph.  This dataflow graph is then iteratively processed until all of the implementation choices have been made.  The compilation process consists of three phases: analysis, lowering, and optimization, which are together iterated until all of the implementation decisions have been made.  The graph is then translated into C code.

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/compile.png}

\caption{Overview of the MATLAB compiler.}
\label{fig:compiler}
\end{figure}

\subsubsection{The Dataflow Graph}

An example dataflow graph for the GEMVER kernel, defined below, is given in figure~\ref{fig:gemver-dataflow}.
\begin{align*}
  A &\gets A + u_1 v_1^T + u_2 v_2^T \\[-0.5ex]
  x &\gets \beta A^T y + z \\[-0.5ex]
  w &\gets \alpha A x
\end{align*}
Each node represents a parameter of the kernel or an operation.  The arrows indicate the flow of data. At first the graph specifies what operations are to be performed but does not contain any implementation details. The symbol $\times$ in the depicted graph stands variously for outer product, matrix-vector multiplication, and scalar-vector multiplication, and it does not yet specify, for example, whether the outer products compute a row or a column at a time of the result matrix.

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/gemver-dataflow.png}

\caption{Dataflow graph for the GEMVER kernel.}
\label{fig:gemver-dataflow}
\end{figure}

\subsubsection{Analyze the Dataflow Graph}

At the beginning of the analysis phase, each input and output node of the dataflow graph is labeled with its kind (scalar, vector, matrix, or unknown) and storage format (dense row-major, dense column-major, or unknown) according to the annotation given by the user.  During analysis, the kind and storage format are computed for the intermediate nodes and algorithms are assigned to operation nodes. The choice of algorithm and the determination of kinds of storage formats are interrelated so they must be computed together.  For example, the $\times$ symbol in the graph in Figure~\ref{fig:gemver-dataflow} having the inputs $u_1$ and $v_1^T$ can be implemented by iterating over rows first, or over columns first. The choice depends on how the result is used downstream in the dataflow graph.  In this case, the result is added to the outer product of $u_2$ and $v_2^T$, so we still could choose either option as long as we make the same choice for both outer products. Going one more step downstream, there is an addition with $A$, which was annotated in Figure~\ref{fig:gemver} to be column major. At this point it is clear that the outer-products should be computed in column-major order.

We do not hard code this knowledge of basic linear algebra in the compiler algorithm itself, but instead use a data-driven approach and store information regarding how to implement basic linear algebra in a separate database, which we refer to as the \emph{linear algebra database}.  This separation allows us to add new matrix formats, operations, and basic linear algebra algorithms without changes to the
compiler algorithm.

The analysis algorithm makes implementation choices using the most-constrained first strategy (also called minimum remaining values)
from the artificial intelligence literature~\cite{Russell:2003mz}.This may need to be updated depending on how we decide to talk about the empirical search. -Jeremy 8/6/08 10:15 AM 
The compiler chooses the node with the fewest matching implementations (in the linear algebra database) and assigns a matching implementation to the node.  If there is more than one match, the prototype compiler picks the first.  Once we add backtracking we will push each alternative onto the stack to be explored later. Once the choice is made, the kind and storage format labels for the operation node itself and for the input nodes are updated with the labels specified in the linear algebra database for the chosen algorithm. The algorithm then goes on to find the next node with the fewest matching implementations and repeats.  In this phase, only the name of the algorithm is assigned to the operation node.  The details of the implementation are not specified until the refinement phase.

\subsection{Refine the Dataflow Graph}
\label{sec:refine}

The refinement phase resolves the implementation for each operation node in the graph into a subgraph defining the details of the chosen algorithm.  Each subgraph is an abstract representation of the loop that implements the given operation.  In each case, the subgraph has an iteration strategy that specifies how to traverse the elements of the matrix or vector.

Figure~\ref{fig:refine-mv-dot} shows two steps of refinement for a matrix-vector multiplication. The first step expands the matrix-vector multiplication according to the mv-dot algorithm from figure~\ref{fig:algos}.  The refinement step replaces the $\times$ node with a subgraph that computes the inner product of a row of the matrix---the node labeled $(i,:)$---with the vector, storing the result in the $i$th element of the result vector---the node labeled $(i)$. The subgraph is labeled with $i=1\!:\!m$ indicating that the iteration strategy is to traverse the rows of the dense matrix.
%
The second pass of refinement introduces another subgraph to implement the inner products (the dot algorithm from figure~\ref{fig:algos}). The new subgraph iterates through each dense row, as indicated by the $j=1:n$ annotation. The sign $\Sigma$ indicates that the sum of all of the iterations is computed.

\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth]{figures/refine-mv-dot.png}

  \caption{The refinement of a matrix-vector product to a set of inner products and the refinement of inner products to scalar multiplications and additions.}
  \label{fig:refine-mv-dot}
\end{figure}

\subsubsection{Optimize the Dataflow Graph}

In the optimization step, the dataflow graph is optimized by applying conditional graph rewrite rules.  One example rewrite rule is \emph{Merge Operand Sharing Subgraphs}: if two subgraphs share a common operand but do not depend on one another, and if the iteration strategies of the two subgraphs are compatible, then merge the subgraphs into a single subgraph.  This rule is responsible for fusing the loops of the two matrix-vector products in the GEMVER kernel.  Another example rule is \emph{Merge Dependent Subgraphs}: if one subgraph depends on another subgraph, and the iteration strategies of the two subgraphs are compatible, then merge them into a single subgraph.  This rule, for example, is responsible for fusing the loops of the two outer products in the GEMVER kernel.

The choice of which optimizations to apply to various parts of the graph depends on the characteristics of the kernel, the algorithm, the matrix order and kind, and architectural features.  The following is an instructive example: $y \gets A^T A (x + z + w)$.  Assume that the compiler has already merged the two vector additions into one subgraph, call it $S_1$, and the two matrix products into another subgraph, call it $S_2$. The question then is whether it is profitable to apply the Merge Dependent Subgraphs rule one more time to merge $S_1$ and $S_2$. If the three vectors plus a row of $A$ do not fit in cache, then the merger will not be profitable because it will cause each row of $A$ to fall out of cache before the row gets used a second time.  On the other hand, if all three vectors plus a row of $A$ do it in cache, then $S_1$ and $S_2$ should be merged.  In general, we need a way to search through the many different combinations of optimization choices and choose a combination that minimizes memory traffic.

\subsubsection{Translate the Dataflow Graph to C}

A graph that cannot be further refined has been reduced to a collection of subgraphs representing loops over scalar operations. The generator outputs a C loop for each subgraph.  The order in which the loops, and scalar operations within the loops, appear in the output is determined by performing a topological sort
\subsection{Orio}
\label{sec:orio}

Orio~\cite{Norris:2007} is an empirical tuning tool that takes as input annotated C code and generates multiple tuned versions of the annotated code. Orio does not parse the application code; rather, it relies on annotations expressed as structured comments, which contain a simplified expression of the computation, as well as directives on what transformations to apply.

% Explain how Pluto is used

% Forward ref. to search approaches

\begin{figure}[htbp]
\centering
\includegraphics[width=.7\textwidth]{figures/orio.png}
\caption{Overview of the Orio empirical tuning process.}
\label{fig:orio}
\end{figure}

\subsection{Search Space Exploration and Evaluation}
\label{sec:search}

This will need updating depending on how we decide to address this topic. -Jeremy 8/6/08 10:19 AM
% BN: moved to its own section since multiple tools deal with the search space problem.
The set of all possible combinations of implementation and optimization choices is called the search space~\cite{Kisuki:2000uq,Triantafyllis:2003uq, Cooper:2005kx}.  Although the dataflow graphs are quite small in our setting (usually less than 100 nodes), it is still important to prune the search space to avoid an exponential compilation time. Our prototype Matlab compiler already performs some pruning: the optimization rewrite rules only apply under certain circumstances. 
%
% BN: commented out below since it's more future work
%However, as mentioned above we are sometimes overly aggressive in applying optimizations.
%To solve this problem we plan to use a combination of analytic methods and empirical methods~\cite{Chen:2007vn}.
%
Orio also offers a number of different general strategies for narrowing the search space over loop-level optimizations, including random, simplex, and simulated annealing searches.

\section{Experimental Results}
\label{sec:experiments}

In this section we present performance results for a number of composed BLAS operations. We performed experiments on an Intel Xeon 2.80 GHz Intel(R) E5462 dual quad-core workstation with 16GB DDR2 FBDIMM RAM.
% Include peak numbers?

\subsection{VADD}

$x = w + y + z$

BN: working on this

\subsection{ATAX}

The extended Matlab code for the ATAX operation, $y = A' * (A * x)$,
and the corresponding performance results are shown in Figure~\ref{fig:gesummv}.

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
ATAX
in
  A : row matrix, x : vector
out
  y : vector
{
  y = A' * (A * x)
}

\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/atax.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the ATAX operation using different code versions.}
\label{fig:atax}
\end{figure}

\subsection{GEMVER}

The extended Matlab code for the GEMVER operation and the corresponding performance results are shown in Figure~\ref{fig:gemver}.

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
GEMVER
in
  A : column matrix,
  u1 : vector, u2 : vector,
  v1 : vector, v2 : vector,
  a : scalar, b : scalar,
  y : vector, z : vector
out
  B : column matrix, 
  x : vector, w : vector
{
  B = A + u1 * v1' + u2 * v2'
  x = b * (B' * y) + z
  w = a * (B * x)
}
\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/gemver.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the GEMVER operation using different code versions.}
\label{fig:gesummv}
\end{figure}

% Data (Dual quad-core Intel Xeon E5462, 16GB RAM)
%--- Sequential: seconds (dynamic arrays) ---
%cmatlab= [0.038, 0.157, 0.352, 0.626, 0.976, 1.406, 1.909, 2.494, 3.143, 3.886]
%orio= [0.029, 0.118, 0.263, 0.480, 0.759, 1.075, 1.447, 1.899, 2.338, 2.978]
%blas= [0.064, 0.265, 0.598, 1.091, 1.711, 2.456, 3.340, 4.345, 5.480, 6.840]
%mkl=  [0.073, 0.147, 0.331, 0.591, 0.923, 1.326, 1.803, 2.350, 2.975, 3.676]
%atlas=[0.039, 0.163, 0.401, 0.796, 1.225, 1.916, 2.582, 3.381, 4.210, 4.756]
%
%--- Parallel: seconds (dynamic arrays) --- N=10,000, cores=1-8
%cmatlab= [0.976, 0.976, 0.976, 0.976, 0.976, 0.976, 0.975, 0.976]
%orio= [0.754, 0.738, 0.666, 0.649, 0.664, 0.664, 0.647, 0.668]

\subsection{GESUMMV}

The extended Matlab code for the GESUMMV operation, $y = a * (A * x) + b * (B * x)$,
and the corresponding performance results are shown in Figure~\ref{fig:gesummv}.

\begin{figure}[htp]
\centering
\begin{tabular}{cc}
\begin{minipage}[b]{.36\textwidth}
\footnotesize
\begin{verbatim}
GESUMMV
in
  A : row matrix,
  B : row matrix,
  x : vector,
  a : scalar,
  b : scalar
out
  y : vector
{
  y = a * (A * x) + b * (B * x)
}

\end{verbatim}
\end{minipage}
&
\begin{minipage}[b]{.6\textwidth}
\includegraphics[width=\textwidth]{figures/gesummv.png}
\end{minipage}\\
\end{tabular}
\caption{Annotated MATLAB code (left) and wall-clock time in seconds (right) for the GESUMMV operation using different code versions.}
\label{fig:gesummv}
\end{figure}

% BN: ATLAS's dgemv is annihilating everything else, even though it's called twice... wow.
% I did check the result and it was correct
% The Time and MFLOPS data for the above are (with icc):
% (Dual quad-core Intel Xeon E5462, 16GB RAM)
% ops = (4*N*N+3*N)
% --- Parallel: seconds (dynamic arrays) ---
%matlabc= [0.019, 0.075, 0.170, 0.301, 0.471, 0.678, 0.923, 1.205, 1.529, 1.887]
%blas= [0.00006, 0.00018, 0.00039, 0.00060, 0.00090, 0.00126, 0.00168, 0.00216, 0.00270, 0.00330]
%orio= [0.013, 0.035, 0.076, 0.133, 0.210, 0.300, 0.410, 0.533, 0.676, 0.831]
%--- Parallel: Mflops/sec (dynamic arrays) ---
%matlabc= [846.565, 848.425, 849.028, 849.503, 849.690, 849.968, 849.605, 849.946, 847.893, 848.023]
%blas= [271288.136, 347891.304, 367392.857, 421785.832, 480228.091, 497440.415, 508787.800, 516153.226, 524930.741, 530172.300]
%orio= [1231.989, 1846.163, 1885.595, 1918.746, 1907.976, 1920.818, 1910.867, 1920.502, 1915.970, 1924.623]
%
% and gcc behaves similarly, so it's not icc being smart and detecting 2 calls to dgemv
% time: blas= [0.000, 0.000, 0.000, 0.001, 0.001, 0.001, 0.002, 0.002, 0.003, 0.003]
% mflops: blas= [254063.492, 346010.811, 393491.803, 421785.832, 439593.407, 453214.792, 463108.092, 470394.120, 462381.020, 469088.244]

\subsection{BiCG Kernel}

$q = A * p$, $s = A' * r$

\begin{figure}[htp]
\centering

\includegraphics[width=.8\textwidth]{figures/bicgkernel.png}
\caption{Wall-clock time (in seconds) for the BiCG Kernel operation using different code versions.}
\label{fig:bicgkernel}
\end{figure}

%NOTES about the experiments

%Results already obtained:
% -gemver (orio/testsuite/gemver/cookie directory)
% -gesummv (orio/testsuite/gesummv/cookie directory)
% -bicgkernel (orio/testsuite/bicgkernel/cookie directory)
% -atax (orio/testsuite/atax/cookie directory)

%* All results were obtained on the cookie machine (eight-core Intel Xeon)

%* Compiler used: icc 10.1
%* Optimization flags used: -O3 (-parallel for parallel matlab code) (-openmp for parallel orio code)

%* We compare two types of codes:
%  1. matlab
%  2. orio
%1. 'matlab' code is the original code produced by the matlab c compiler.
%2. 'orio' is obtained first from the Pluto-generated code (use loop fusion, NO loop tiling,
%   NO unrolling, NO parallelization, NO vectorization), then was decorated with
%   performance annotations before being tuned by Orio.

%* Each code has two versions: one with static arrays, and another with dynamic arrays.

%* Conversion from/to static arrays to/from dynamic arrays was done manually.

%* I run Pluto first to maximally fuse all loops (using polycc <file>
%--maxfuse option) and see the resulting fused loop(s). And then tune the
%fused loop using Orio. Sometimes, the fused loop gets very complicated
%because of Pluto's various transformations (e.g. skewing). In that case, I
%don't use Pluto's loop fusion. For example, gemver and gesummv results use
%Pluto's loop fusion, but bicgkernel doesn't use Pluto's loop fusion at
%all.

%* All experiment files are in <kernel-name>/cookie directory

%* All generated codes (the base Matlab codes, and the Orio-tuned codes),
%results, script for running the experiments are in
%<kernel-name>/cookie/exp

%* I tuned 4 different versions for each kernel:
% 1. sequential (static arrays)
% 2. sequential (dynamic arrays)
% 3. parallel (static arrays)
% 4. parallel (dynamic arrays)

%The annotated source code for 1 and 3 is in <kernel-name>.src1.c
%The annotated source code for 3 and 4 is in <kernel-name>.src2.c

%The final code for 1 and 2 is in <kernel-name>.orio.seq.c
%The final code for 3 and 4 is in <kernel-name>.orio.par.c

%* <kernel-name>.matlab.c contains the original Matlab code (for both
%static and dynamics arrays)

%* The best performance parameters for each version can be seen in
%results.txt file.

\section{Conclusions and Future Work}

fully automate the process to get rid of the manual steps along the way

extend beyond composed BLAS

Ultimately, the tuning process should result in more than one version of the computation since different versions are optimal for different problem sizes.

This motivates further work on automating more parts of the process, so that these tools can be used by application developers to automatically produce custom linear algebra libraries tuned for specific architectures.

% BN: moved from search space section above
%Yotov:2005kx
We plan to use a fast analytic cost function to prune out regions of the search space that will not produce competitive implementations. The cost function is based on a model of the computer architecture, but does not necessarily account for all the details.  Once the search space is narrowed, we will use empirical testing with representative data sets to determine the exact performance of the alternatives and choose the best one.

\singlespacing
\bibliographystyle{siam}
\bibliography{paper}
\end{document}