### 2008-10-13 doc/papers/ipdps09/paper.tex

\documentclass[12pt]{article}
%\usepackage{latex8}
%\usepackage[]{rotating} 
\usepackage{ifpdf}
\usepackage{url} 
\usepackage{comment} 
\usepackage{cite} 
\usepackage{graphicx} 
\usepackage{subfigure}  
\usepackage{wrapfig}
\usepackage{times}
\usepackage{amsmath,amsthm}
\usepackage[letterpaper,margin=1in]{geometry}

\ifpdf
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\else
\DeclareGraphicsExtensions{.eps,.ps}
\fi

\newenvironment{itemizer}{\begin{itemize}\setlength{\parsep}{0cm}\setlength{\itemsep}{-.3em}}{\end{itemize}}

\long\def\comment#1{}
\date{}

\title{\vspace{-.6in}Annotation-Based Empirical Performance Tuning\\ Using Orio}

\author{
Albert Hartono$^1$, Boyana Norris$^2$, and P. Sadayappan$^1$ \\
{\small $^1$Ohio State University, Dept. of Computer Science and Engineering, Columbus, OH 43210}\\
{\small \texttt{\{hartonoa, saday\}@cse.ohio-state.edu}}\\
{\small $^2$Argonne National Laboratory, Mathematics and Computer Science Division, Argonne, IL 60439}\\
{\small \texttt{norris@mcs.anl.gov}}
}

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\input{motivation}
\input{related}
\input{implementation}
\input{integration}
\input{results} 
\input{conclusions}

\small
\bibliographystyle{abbrv} 
\bibliography{paper}

\end{document}