### 2008-10-13 doc/papers/ipdps09/motivation.tex

\section{Motivation}
\label{sec:motivation}

%The use of empirical performance tuning and annotations approaches for
%the following purposes:
%- improve productivity of scientific computation applications
%- improve high performance by (semi-)automating the tuning process
%- improve code readability
%- alleviate performance portability issues

The size and complexity of scientific computations are increasing at least as
fast as the improvements in processor technology.  Programming such
scientific applications is hard, and optimizing them for high performance is
even harder.  This results in a potentially large gap between the achieved
performance of applications and the peak available performance, with
many applications achieving 10\% or less of the peak (e.g.,~
%\cite{Oliker,
\cite{conf/vecpar/CarterOS06}).  
A greater concern is the inability of existing languages, compilers, and
systems to deliver the available performance for the application through
fully automated code optimizations.

Delivering performance without degrading productivity is crucial for the
success of scientific computing. Scientific code developers generally attempt
to improve performance by applying one or more of the following three
approaches: manually optimizing code fragments, using tuned libraries for key
numerical algorithms, and, less frequently, using compiler-based source
transformation tools for loop-level optimizations. Manual tuning is
time-consuming and impedes readability and performance portability. Tuned
libraries often deliver great performance without requiring significant
programming effort, but can only provide limited
functionality. General-purpose source transformation tools for performance
optimizations are few and have not yet gained popularity among computational
scientists, at least in part because of poor portability and steep learning
curves.