### 2009-02-12 doc/papers/ipdps09/paper.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@428 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

\documentclass[conference]{IEEEtran}

%\documentclass[12pt]{article}
%\usepackage{latex8}
%\usepackage[]{rotating} 
\usepackage{ifpdf}
\usepackage{url} 
\usepackage{comment} 
\usepackage{cite} 
\usepackage{graphicx} 
\usepackage{subfigure}  
\usepackage{wrapfig}
\usepackage{times}
\usepackage{amsmath,amsthm}

%\usepackage[letterpaper,margin=1in]{geometry}

%\ifpdf
%\DeclareGraphicsExtensions{.pdf,.png,.jpg}
%\else
%\DeclareGraphicsExtensions{.eps,.ps}
%\fi

%\newenvironment{itemizer}{\begin{itemize}\setlength{\parsep}{0cm}\setlength{\itemsep}{-.3em}}{\end{itemize}}

\long\def\comment#1{}
\date{}

\begin{document}

\title{Annotation-Based Empirical Performance Tuning Using Orio}

\author{
\IEEEauthorblockN{Albert Hartono} 
\IEEEauthorblockA{\small Dept. of Computer Science and Engg.\\ 
Ohio State University\\ 
Columbus, Ohio 43210--1277\\ 
Email: hartonoa@cse.ohio-state.edu} 
\and 
\IEEEauthorblockN{Boyana Norris} 
\IEEEauthorblockA{\small Mathematics and Computer Science Division\\ 
Argonne National Laboratory\\ 
Argonne, Illinois 60439--4844\\
Email: norris@mcs.anl.gov} 
\and 
\IEEEauthorblockN{P. Sadayappan} 
\IEEEauthorblockA{\small Dept. of Computer Science and Engg.\\ 
Ohio State University\\ 
Columbus, Ohio 43210--1277\\ 
Email: saday@cse.ohio-state.edu}
}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\IEEEpeerreviewmaketitle

\input{motivation}
\input{related}
\input{implementation}
\input{integration}
\input{results} 
\input{conclusions}

\section*{Acknowledgment} 
The authors would like to thank Uday Bondhugula of IBM Research for
many productive discussions and his valuable help with Pluto. This
work was supported in part by the U.S. Dept. of Energy under Contract
DE-AC02-06CH11357.

\small
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,paper}

\end{document}