### 2013-11-03 doc/papers/ipdps09/abstract.tex

For many scientific applications, significant time is spent in tuning
codes for a particular high-performance architecture. Tuning
approaches range from the relatively nonintrusive (e.g., by using
compiler options) to extensive code modifications that attempt to
exploit specific architecture features. Intrusive techniques often
result in code changes that are not easily reversible, and can
negatively impact readability, maintainability, and performance on
different architectures. We introduce an extensible annotation-based
empirical tuning system called Orio that is aimed at improving both
performance and productivity. It allows software developers to insert
annotations in the form of structured comments into their source code
to trigger a number of low-level performance optimizations on a
specified code fragment.  To maximize the performance tuning
opportunities, the annotation processing infrastructure is designed to
support both architecture-independent and architecture-specific code
optimizations. Given the annotated code as input, Orio generates many
tuned versions of the same operation and empirically evaluates the
alternatives to select the best performing version for production
use. We have also enabled the use of the Pluto automatic
parallelization tool in conjunction with Orio to generate efficient
OpenMP-based parallel code. We describe our experimental results
involving a number of computational kernels, including dense array and
sparse matrix operations.