### 2013-11-03 doc/papers/ipdps09/motivation.tex

\section{Motivation}
\label{sec:motivation}

%The use of empirical performance tuning and annotations approaches for
%the following purposes:
%- improve productivity of scientific computation applications
%- improve high performance by (semi-)automating the tuning process
%- improve code readability
%- alleviate performance portability issues

The size and complexity of scientific computations are increasing at
least as fast as the improvements in processor technology.
Programming such scientific applications is hard, and optimizing them
for high performance is even harder.  This results in a potentially
large gap between the achieved performance of applications and the
peak available performance, with many applications achieving 10\% or
less of the peak (e.g.,~\cite{Oliker,conf/vecpar/CarterOS06}).  A
greater concern is the inability of existing languages, compilers, and
systems to deliver the available performance for the application
through fully automated code optimizations.

Delivering performance without degrading productivity is crucial for
the success of scientific computing. Scientific code developers
generally attempt to improve performance by applying one or more of
the following three approaches: manually optimizing code fragments,
using tuned libraries for key numerical algorithms, and, less
frequently, using compiler-based source transformation tools for
loop-level optimizations. Manual tuning is time-consuming and impedes
readability and performance portability. Tuned libraries often deliver
excellent performance without requiring significant programming effort,
but can only provide limited functionality. General-purpose source
transformation tools for performance optimizations are few and have
not yet gained popularity among computational scientists, at least in
part because of poor portability and steep learning curves.