### 2013-11-05 doc/papers/ipdps09/related.tex

\section{Related Work}
\label{sec:related}

%Optimization abstraction:
%- POET (Qing Yi)
%- Apan Qasem's thesis work (Ken Kennedy's student)
%- ...
%
%Automated performance tuning:
%- ATLAS
%- SPIRAL
%- FFTW
%-...

Ideally, a developer should only have to specify a few simple
command-line options and then rely on the compiler to optimize the
performance of an application on any architecture. Compilers alone,
however, cannot fully satisfy the performance needs of scientific
applications.  First, compilers must operate in a black-box fashion
and at a very low level, limiting both the type and number of
optimizations that can be done.  Second, static analysis of
general-purpose languages, such as C, C++, and Fortran, is necessarily
conservative, thereby precluding many possible optimizations.  Third,
in the process of transforming a mathematical model into a computer
program, much potentially useful (for optimization purposes)
information is lost since it cannot be represented by the programming
language.  Finally, extensive manual tuning of a code may prevent
certain compiler optimizations and result in worse performance on new
architectures, resulting in loss of performance portability.

%As briefly discussed in Section~\ref{sec:motivation}, performance tuning is
%generally approached in three ways: by performing manual optimizations of key
%portions of the code; by using compiler-based source transformation tools for
%loop optimizations; and by using tuned libraries for key numerical
%algorithms.
%% Libraries

An alternative to manual or automated tuning of application codes is
the use of tuned libraries. The two basic approaches to supplying
high-performance libraries include providing a library of hand-coded
options (e.g., \cite{BLAS,ESSL,Goto:2006fk}) and generating optimized
code automatically for the given problem and machine
parameters. ATLAS~\cite{atlas_sc98,WN147} for BLAS~\cite{BLAS} and
some LAPACK~\cite{laug} routines, OSKI~\cite{OSKI} for sparse linear
algebra, PHiPAC~\cite{bilmes97optimizing} for matrix-matrix products,
and domain-specific libraries such as FFTW~\cite{frigo98} and
SPIRAL~\cite{SPIRAL} are all examples of the latter approach. Most
automatic tuning approaches perform empirical parameter searches on
the target platform.  FFTW uses a combination of static models and
empirical techniques to optimize FFTs. SPIRAL generates optimized
digital signal processing libraries by an extensive empirical search
over implementation variants.  GotoBLAS~\cite{Goto:2006fk,Goto:fk}, on
the other hand, achieves near-peak performance on several
architectures by using hand-tuned data structures and kernel
operations.  These auto- or hand-tuned approaches can deliver
performance that can be five times as fast as that produced by many
optimizing compilers
\cite{WN147}.  The library approach, however, is limited by the fact
that optimizations are highly problem- and
machine-dependent. Furthermore, at this time, the functionality of the
currently available automated tuning systems is quite limited.

%% Other annotation-based source transformation approaches

General-purpose tools for optimizing loop performance are also
available.  LoopTool~\cite{LoopTool} supports annotation-based loop
fusion, unroll/jamming, skewing and tiling.  The Matrix Template
Library \cite{Siek:1998ys} uses template metaprograms to tile at both
the register and cache levels.  A new tool, POET~\cite{POET} also
supports a number of loop transformations. POET offers a complex
template-based syntax for defining transformations in a
language-independent manner. Other research efforts whose goal, at
least in part, is to enable optimizations of source code to be
augmented with performance-related information include the X
language~\cite{XLanguage} (a macro C-like language for annotating C
code), the Broadway~\cite{broadway} compiler, and telescoping
languages~\cite{telescopingurl,teleoverview,Ken99}, and various
meta-programming
techniques~\cite{veldhuizen95,weise93,kiczales91,chiba95}.

Emerging annotation-based tools are normally designed by compiler researchers
and thus the interfaces are not necessarily based on concepts accessible to
computational scientists. The complexity of existing annotation languages and lack
of common syntax for transformations (e.g., loop unrolling) result
in steep learning curves and the inability to take advantage of more than one
approach at a time. Furthermore, at present, there is no good way for
users to learn about the tools available and compare their
capabilities and performance.